---
title: "Stickbreaker Vignette v2"
author: "Craig Miller and James Van Leuven"
date: "`r date()`"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 5
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 5
csl: apa.csl  
---

# I. Introduction {-}

Genetically modified organisms are used to test the phenotypic effects of mutation. Often, multiple substitutions impact organismal fitness differently than single substitutions (epistatic interactions). This package fits three basic models (additive, multiplicative, and stickbreaking) to fitness data and suggests the best fitting model by multinomial regression. Stickbreaker can also be used to simulate fitness data.

```{r setup, echo=FALSE, include=FALSE}
suppressPackageStartupMessages({
  library(nnet)
  library(lattice)
  library(xtable)
  devtools::load_all()
})
```

# II. Basic Analysis {-}

## A. Data format {-}

The data should be in a text file. If there are $n$ mutations in the dataset, there should be $n$ columns in the file. The first row should be mutation names. The data in the columns are 0/1 to indicate the absence/presence of the mutation. All 0s indicates wild type and should be the first row of data. The last column is the estimated fitness of that genotype. You can see the format of the data below.

## B. Read in data {-}


```{r, echo=TRUE}
  inpath <- system.file("data", package="Stickbreaker")
  data <- read.csv(paste(inpath,"Chou.data.csv",sep="/"))
```

## C. Fit models {-}

The `fit.models()` function is a wrapper that fits the data to all three models. First, it estimates d and, using this, fits the stickbreaking model. Then it fits the multipicative and additive models. If you want to do this fitting to models individually and better understand what is going on, see the Detailed Features section.

To fit the models, you need to specify a few things. `d.range` specifies the range of possible values for d. If you fitness values are unusually large you may want to change this range. The upper bound is more important. The reason not to set this really large (say to 100 or 1000) is that the stickbreaking model approaches the additive model as d gets large and the coefficients get small. By putting a modest upper bound on stickbreaking, we demand that additivity remains a distinct model.

When the stickbreaking model doesn't fit the data very well or when the data is really noise, the preffered ways to estimate d can fail. Still, we need to fit the stickbrekaing model (albiet poorly) for comparision reason. We do this simply by multiplying the fitness distance between the wildtype and the largest observed fitness by a factor, `d.adj.max`. The default is 1.1. By making it only slightly above one, we ensure that stickbreaking remains distinct from the additive model.

`wts` specifies how to weight single mutations (genotypes on the wildtype background) vs all other genotypes when estimating the coefficients. If the wildtype fitness is know which much greater accuracy than the other genotypes, then the comparisons to it will have lower variance than the other comparisions. The default weighting is `c(2,1)` which indicates to weight the single mutations twice as heavily as all others which corresponds to our assumption that the wildtype fitness is know without error while all other genotypes have the same error.


```{r, echo=TRUE}
  fit.results <- fit.models(data, d.range=c(0.1, 10), d.adj.max=1.1, wts=c(2,1))
```

`fit.models()` returns a list of results (saved here as `fit.results`).  `fit.results$fit.smry` is the most important item. This is a vector of $R^2$ and P-scores--one each from the three models--that summarize model fits and to be used in the next code block. These summary statistics will be used to determine the posterior probability of each. See paper for details about how these fit statistics are calculated. \cr

The other items in the list are details of fit to each model: `$fit.stick`, `$fit.mult` and `$fit.add`. Each of these contains analagous information, itself organized as a list. [[1]] The coefficients (`$u.hats` for stickbreaking, `$s.hats` for multiplicative, and `$w.hats` for additive). [[2]] `$R2`.  [[3]] `$sig.hat` is estimate of $\sigma$. [[4]] `$log.like`.  [[5]] `$regression.results` is itself a list with `$p.vals`, `lm.intercepts` and `lm.slopes` being the linear regression p-values, intercepts, and slopes for each mutation and `P` being the P-score. The last two items of `$regression.results` are matrices `$fitness.of.backs` and `$effects.matrix`. These are the data being regressed.  `$fitness.of.backs` gives the fitness of the background genotype when a given mutation (column) is added to createa given genotype (row). `$effects.matrix` gives the fitness effect that this mutation has when added in the coefficient of the model considered.\cr

## D. Generate multinomial and calculate posterior probabilities {-}

After the data has been fit to the three models, we use the summary fit statistics to assign posterior probabilities to the three models. This is done in three steps: (1) simulate data from priors, (2) fit simulated data to multinomical regression and (3) calcualte posterior probabilities.\cr

*Priors etc.:* To simulate data you need to define your priors and the number of replicate datasets to simulate. These are defined at the beginning of the next code block. `coes.prior` is a vector with the lower and upper bounds on the coefficients; `sig.prior` is the lower and upper bound on the $\sigma$. `n.samps.per.mod` is the number of datasets to simulate per model. You may want to initially set this to a small value like 50 or 100 to check things are running and behaving properly. For a final analysis you should increase this number to 1000 or better yet 10000. Be warned that it will take some time to simulate 10000 datasets for each model. `d.range`, `d.adj.max`, and `wts` are discussed above. Finally, use `analysis.name` to assign a unique name so that the simulated fitted data and the multinomial model it leads to are saved with unique identifiers. These files may be of interest (and save you computing time) if you choose to tap into any of the detailed features detailed below.

Add mention of `min.R2`.


```{r, echo=TRUE}
  coes.prior <- c(0.05, 0.5)
  sig.prior <- c(0, 0.25)
  n.samps.per.mod <- 25
  d.range <- c(0.1, 10)
  d.adj.max <- 1.1
  wts <- c(2,1)
  min.R2 <- -1
  analysis.name <- "Test"
    
  posterior.results <- simulate.data.calculate.posteriors(fit.results$fit.smry, data, analysis.name, coes.prior, sig.prior, d.range, d.adj.max, wts, n.samps.per.mod, min.R2)
```

A small change
Another change
Another change - jt


# III. Detailed Features {-}

## A. Setup {-}

Whereas above we used wrapper functions to do the entire analysis is just a few commands, here we walk through the various functions in greater deatil.

Read in data. See above for format of data.

```{r}
  inpath <- system.file("data", package="Stickbreaker")
  data <- read.csv(paste(inpath,"Khan.data4.csv",sep="/"))
```

Define parameters to use in estimation (`wts`, `d.range` and `d.adj.max`). Again, these are discussed above. Then extract number of genotypes (`n.genos`), number of mutations (`n.muts`), the genotype matrix (`geno.matrix`; the 0/1 columns in the data) and the fitness matrix (`fit.matrix`; the last column in the data). 

```{r}
  wts <- c(2,1)
  d.range <- c(0.1, 10)
  d.adj.max <- 1.1
  
  n.genos <- length(data[,1])
  n.muts <- length(data[1,])-1
  geno.matrix <- data[,seq(1, n.muts)]
  fit.matrix <- as.matrix(data[,(n.muts+1)])
```

## B. Model fitting {-}

### 1. Fit to stickbreaking model {-}
We have two good estimators of the distance to the fitness boundary. We call them using the functions `d.hat.MLE()` and `d.hat.RDB()`. We take the MLE as the estimate unless it fails; in which case we use the RDB estimate. If both fail, we simply use the distance from wild type to the largest observed fitness value adjusted upwards by a factor `d.adj.max`. This is done in `d.hat.seq`. Why adjust the maximum estimate upwards at all?  If we do not, the genotype with maximum fitness is at the boundary which is problematic since it results in a zero in the denominator when estimating d for this genotype. 

```{r}
  d.hat.MLE <- estimate.d.MLE(geno.matrix, fit.matrix, d.range=d.range)
  d.hat.RDB <- estimate.d.RDB(geno.matrix, fit.matrix)$d.hat.RDB
  d.hat.seq <- estimate.d.sequential(geno.matrix, fit.matrix, 
                                     d.hat.MLE, d.hat.RDB, d.range, d.adj.max)
  d.hat.MLE
  d.hat.RDB
  d.hat.seq
```

After getting an estimate for d, we call `fit.stick.model.given.d()`. This returns a list the  with the stickbreaking coefficients (`$u.hats`), $R^2$ (`$R2`), $\hat{\sigma}$ (`$sig.hat`) and the log-likelihood (`$log.like`). \cr

The function also performs a linear regression of background fitness against effect. When data are generated and analyzed under the same model, the slope of this regression line is expected to be zero for each mutation with p-values distributed U(0,1). When the data are generated and analyzed under different models, the regression line is expected to be non-zero. One regression is performed per mutation. In each case we calcualte the p-value under the null (no slope). We then summarize the data across mutations by taking the sum of the logs of the p-values and call this sum the P-score. The more negative the p-score, the stronger the evidence against a model being correct. The regression results are returned under the list item `$regression.results` with the sub-items giving p-values (`$p.vals`), intercepts (`$lm.intercepts`), and slopes (`$lm.slopes`) and the overall P-score (`$P`).   The last two items of `$regression.results` are matrices `$fitness.of.backs` and `$effects.matrix`. These are the data being regressed.  `$fitness.of.backs` gives the fitness of the background genotype when a given mutation (column) is added to createa given genotype (row). `$effects.matrix` gives the fitness effect that this mutation has when added in the coefficient of the model considered.\cr

Finally, a matrix is returned (`$pred.matrix`) that contains model predcited fitness for each genotypes (appended to the data itself along with error and a binary string that can be useful in producing plots).

```{r}
  fit.stick <- fit.stick.model.given.d(geno.matrix, fit.matrix, d.hat.seq, run.regression=TRUE)
  #fit.stick
```


### 2. Fit dataset to multiplicative model {-}

Here we fit to the multiplicative model using `fit.mult.model()`. The output is exactly analagous to that returned from `fit.stick.model()` discussed above except the coefficients are named `$s.hats` (instead of `u.hats`).
```{r}
  fit.mult <- fit.mult.model(geno.matrix, fit.matrix, wts=wts)
  #fit.mult
```


### 3. Fit dataset to additive model {-}

And finally we fit the additive model using `fit.add.model()`. Again the outputs are the same except coefficients are `$w.hats`.

```{r}
  fit.add <- fit.add.model(geno.matrix, fit.matrix, wts=wts)
  #fit.add
```

### 4. Visualize model fit {-}

Let's visualize the regression of background fitness vs effect size. This should help you see what is going on under the hood.

```{r}
  outdir <- system.file("figures", package="Stickbreaker")
  plot.name <- "Ex_fit_vs_effect_plot.svg"
  file.path <- paste(outdir, plot.name, sep="/")
  svg(file=file.path, width=8, height=8)

  layout(mat=matrix(nrow=4, ncol=(1+n.muts), data=seq(1,4*(1+n.muts)), byrow=T), widths=c(0.25, rep(3,n.muts)), heights=c(0.25, rep(3,3))) -> l
  #layout.show(l)
  par(mar=c(0,0,0,0))
  plot(0,type='n',axes=FALSE,ann=FALSE)
  mod.names <- c("STICK", "MULT", "ADD")
  mut.names <- colnames(data)[1:n.muts]
  mod.cols <- c("deepskyblue","red", "yellow")
  for (i in 1:n.muts){
    plot(0,type='n',axes=FALSE,ann=FALSE, ylim=c(0,1), xlim=c(0,1))
    text(0.5, 0.5, labels=mut.names[i], font=2)
  }
  #--- STICK ---
  par(mar=c(0,0,0,0))
  plot(0,type='n',axes=FALSE,ann=FALSE, ylim=c(0,1), xlim=c(0,1))
  text(0.5, 0.5, labels=mod.names[1], font=2, srt=90)
  par(mar=c(4,4,1,1))
  for (mut.i in 1:n.muts){
    plot(x=fit.stick$regression.results$fitness.of.backs[,mut.i], y=fit.stick$regression.results$effects.matrix[,mut.i], ylab="Effect", xlab="Back fitness", pch=21, bg=mod.cols[1])
    abline(fit.stick$regression.results$lm.intercepts[mut.i], fit.stick$regression.results$lm.slopes[mut.i])
  }
  #--- MULT ---
  par(mar=c(0,0,0,0))
  plot(0,type='n',axes=FALSE,ann=FALSE, ylim=c(0,1), xlim=c(0,1))
  text(0.5, 0.5, labels=mod.names[2], font=2, srt=90)
  par(mar=c(4,4,1,1))
  for (mut.i in 1:n.muts){
    plot(x=fit.mult$regression.results$fitness.of.backs[,mut.i], y=fit.mult$regression.results$effects.matrix[,mut.i], ylab="Effect", xlab="Back fitness", pch=21, bg=mod.cols[2])
    abline(fit.mult$regression.results$lm.intercepts[mut.i], fit.mult$regression.results$lm.slopes[mut.i])
  }
  #--- ADD ---
  par(mar=c(0,0,0,0))
  plot(0,type='n',axes=FALSE,ann=FALSE, ylim=c(0,1), xlim=c(0,1))
  text(0.5, 0.5, labels=mod.names[2], font=2, srt=90)
  par(mar=c(4,4,1,1))
  for (mut.i in 1:n.muts){
    plot(x=fit.mult$regression.results$fitness.of.backs[,mut.i], y=fit.mult$regression.results$effects.matrix[,mut.i], ylab="Effect", xlab="Back fitness", pch=21, bg=mod.cols[2])
    abline(fit.mult$regression.results$lm.intercepts[mut.i], fit.mult$regression.results$lm.slopes[mut.i])
  }
  dev.off()
  
```

Plot the figure just created.

```{r, out.width = "800px"}
    knitr::include_graphics(file.path)
```

Now, let's look at the observed vs model predicted values. This is the other way we assess the fit of the model.

```{r}
  outdir <- system.file("figures", package="Stickbreaker")
  plot.name <- "Ex_pred_obs_fit.svg"
  file.path <- paste(outdir, plot.name, sep="/")
  svg(file=file.path, width=8, height=6)

  mod.cols <- c("deepskyblue","red", "yellow")
  lims <- c(1,1.4)
  par(mar=c(5,4,4,1))
  plot(x=fit.matrix, y=fit.stick$pred.matrix$pred, ylim=lims, xlim=lims, ylab="Model Predicted Fitness", xlab="Observed Fitness", pch=21, bg=mod.cols[1], cex=1.1)
  abline(0,1, lty="dashed")
  points(x=fit.matrix, y=fit.mult$pred.matrix$pred, pch=21, bg=mod.cols[2], cex=0.9)  
  points(x=fit.matrix, y=fit.add$pred.matrix$pred, pch=21, bg=mod.cols[3], cex=0.8) 
  text(x=fit.matrix, y=fit.stick$pred.matrix$pred, labels=fit.stick$pred.matrix$string, srt=90, cex=0.7, pos=3, off=2)
  legend("topleft", legend=c("Stick", "Mult", "Add"), pch=21, pt.bg=mod.cols, bty="n")
  dev.off()
```

Plot the figure just created.

```{r, out.width = "800px"}
    knitr::include_graphics(file.path)
```


## C. Simulation Machinery {-}

The method itself uses simulations to calculate posterior probabilites. We also did a lot of simulation work in the study itself. We show the intersested user how to use the existing code to simulate datasets.

### 1. Single Datasets {-}

#### a. Stickbreaking {-}
Notice first we set the parameters: number of mutations, stickbreaking coefficients, sigma.
```{r}
  n.muts <- 3
  coe <- 0.1
  coes <- rep(coe, n.muts)
  sigma <- 0.1
  w.wt <- 1         #fitness of wild type
  d.true <- 1       # distance to fitness boundary
  n.genos <- 2^n.muts     # number of genotypes in full network
  d.range <- c(0.1, 10)
  d.adj.max <- 1.1

  geno.matrix <- generate.geno.matrix(n=n.muts)
  stick.sim.data <- simulate.stick.data(n.muts, coes, sigma, d.true, w.wt, geno.matrix)
  fit.matrix <- stick.sim.data$fit.matrix
  #print(fit.matrix)
```

#### b. Multiplicative {-}

```{r}
  n.muts <- 3
  selcoe <- 0.3
  selcoes <- rep(selcoe, n.muts)
  sigma <- 0.1
  w.wt <- 1         #fitness of wild type
  n.genos <- 2^n.muts     # number of genotypes in full network

  geno.matrix <- generate.geno.matrix(n=n.muts)
  mult.sim.data <- simulate.mult.data(n.muts, selcoes, sigma, w.wt, geno.matrix)
  fit.matrix <- mult.sim.data$fit.matrix
  #print(fit.matrix)
```

#### c. Additive {-}

```{r}
  n.muts <- 3
  addcoe <- 0.3
  addcoes <- rep(addcoe, n.muts)
  sigma <- 0.1
  w.wt <- 1         #fitness of wild type
  n.genos <- 2^n.muts     # number of genotypes in full network

  geno.matrix <- generate.geno.matrix(n=n.muts)
  add.sim.data <- simulate.add.data(n.muts, addcoes, sigma, w.wt, geno.matrix)
  fit.matrix <- add.sim.data$fit.matrix
  #print(fit.matrix)
```

### 2. Batch of datasets  {-}

Here we show how the code can be used to generate many datasets under a given model and then fit each of them to that same model. This is good for assessing estimation error and bias. Notice in the next code block that now paramters are defined as vectors: `mut.vals`, `coe.vals`, `sig.vals`.  The code will simulate under all combinations of the parmater values (i.e. `length(mut.vals) * length(coe.vals) * length(sig.vals)`). We also define the place to write the data to here.  If you have already simulated data and want to skip that step, set `simulate.batch.data <- FALSE` at the beginning of the next code block. Also notice we can supress update printing by setting `print.status` to FALSE. Finally, notice that there is a boolean argument called `run.regression`. This is used in model selection (nexts section). Here we are simply showing how to simulate data and, to save computational time, have toggled `run.regression` to FALSE. \cr

#### a. Stickbreaking {-}

The only added complexity for stickbreaking is the `fit.methods` parameter. This refers to what method of estimating d should be used. If set to `All` it will estimate d using each possible method and output the results indivdiually (i.e. MLE, RDB, max, and seq). Or you can select an indivdiual method (e.g. `seq`. Use `?simulate.fit.stick.data.batch` to get details.

```{r}
  simulate.batch.data <- TRUE
  
  mut.vals <- c(3,4,5)
  coe.vals <- c(0.1, 0.3, 0.5)
  sig.vals <- c(0.02, 0.05, 0.08)
  w.wt <- 1
  d.true <- 1
  wts <- c(2,1)
  n.reps.ea <- 5   
  d.range <- c(0.1, 10)
  fit.methods <- "seq"
  d.max.adj <- 1.0
  run.regression <- FALSE  
    
  if (simulate.batch.data == TRUE){
    outfile <- "Stick_batch_out_test"
    outdir <- system.file("extdata", package="Stickbreaker")
    outpath <- paste(outdir, outfile, sep="/")
    simulate.fit.stick.data.batch(mut.vals=mut.vals, coe.vals=coe.vals, sig.vals=sig.vals, d.true=d.true, d.range=d.range, w.wt=w.wt, n.reps.ea=n.reps.ea, print.status=FALSE, fit.methods=fit.methods, outpath=outpath, wts, d.max.adj=d.max.adj, run.regression, RDB.method="pos")
  }
```

#### b. Multiplicative {-}

```{r}
  simulate.batch.data <- TRUE
  
  mut.vals <- c(3,4,5)
  coe.vals <- c(0.1, 0.3, 0.5)
  sig.vals <- c(0.02, 0.05, 0.08)
  w.wt <- 1
  n.reps.ea <- 5
  
  #--- mulitiplicative ---
  if (simulate.batch.data == TRUE){
    outfile <- "Mult_batch_out_test"
    outdir <- system.file("extdata", package="Stickbreaker")
    outpath <- paste(outdir, outfile, sep="/")
    simulate.fit.mult.add.data.batch(epi.model="mult", mut.vals=mut.vals, coe.vals=coe.vals, sig.vals=sig.vals, w.wt=w.wt, n.reps.ea=n.reps.ea, print.status=FALSE, outpath=outpath, wts)
  }
```


#### c. Additive {-}

```{r}
  simulate.batch.data <- TRUE
  
  mut.vals <- c(3,4,5)
  coe.vals <- c(0.1, 0.3, 0.5)
  sig.vals <- c(0.02, 0.05, 0.08)
  w.wt <- 1
  n.reps.ea <- 5
  
  if (simulate.batch.data == TRUE){
    outfile <- "Add_batch_out_test"
    outdir <- system.file("extdata", package="Stickbreaker")
    outpath <- paste(outdir, outfile, sep="/")
    simulate.fit.mult.add.data.batch(epi.model="add", mut.vals=mut.vals, coe.vals=coe.vals, sig.vals=sig.vals, w.wt=w.wt, n.reps.ea=n.reps.ea, print.status=FALSE, outpath=outpath, wts)
  }
```



## D. Model Selection {-}

### 1. Full network data {-}

We begin by focusing on training to full network data (i.e. when all `2^n.muts` genotypes are present. Below we show how to do model selection incomplete networks.

#### a. Simulate training data from priors {-}

To perform model selection, our Bayesian approach is based on simulating data under each model
and under a range of potential parameter values. Each simulated dataset is fit to each of the three models. We simulate training data by drawing parameter values from uniform priors. Toggle \code{simulate.training.data.from.priors} to FALSE if you want to skip this or if you already have the data simulated. The parameter `n.samps.per.mod` defines how many datasets to simulate per model. Once you have the code running smoothly and you want to generate a training dataset to be used in further analysis, set this to large number (e.g. 10000). \cr

*Priors*:  `coes.prior` and `sig.prior` specify the prior range of values to simulate coefficiets and sigma values from. Be sure these are reasonable ranges for the data you are dealing with. Also see Basic Analysis section above about priors and other definitions in this code block such as `d.range`, `d.adj.max`, and `wts`. 




```{r}

  simulate.training.data.from.priors <- TRUE

  if (simulate.training.data.from.priors == TRUE){
    coes.prior <- c(0.05, 0.5)
    sig.prior <- c(0, 0.25)
    n.samps.per.mod <- 10
    d.true <- 1
    d.range <- c(0.1, 10)
    d.adj.max <- 1.1
    w.wt <- 1
    wts <- c(2,1)
    muts.to.sim <- c(3,4,5)
    print.interval <- NA  # set to NA to block update printing
    
    for (mut.i in 1:length(muts.to.sim)){
      n.muts <- muts.to.sim[mut.i]
      print(paste("n.muts=", n.muts))
      outdir <- system.file("extdata", package="Stickbreaker")
      file.name <- paste("Training_simulated_priors_fit_data_", n.muts, "muts.txt", sep="")
      outpath <- paste(outdir, file.name, sep="/")
      simulate.data.from.priors.for.mod.selection(n.muts=n.muts, coes.prior=coes.prior, sigs.prior=sig.prior, mods.to.sim=c("stick", "mult", "add"), d.true=d.true, d.range=d.range, d.adj.max=d.adj.max, w.wt=w.wt, wts=wts, outpath=outpath, n.samps.per.mod=n.samps.per.mod, coe.sim.model="identical", coe.dist.par=NA, print.interval=print.interval)
    } #next mut.i
  }
```

#### b. Fit training data to multinomial model for model selection {-}

Here we pass the training datasets to the `multinom()` function in the `nnet` package. It fits the data to a multinomial regression model and calculates the posterior probabillity of each model. We do this separately for each number of mutations. `fit.nnet.multinomial.regression()` is the function that handles this. The fitted model is then saved in in the directory `inst/models` using the naming convention `nnet_mod_R2_P_nmuts.rda`. If you want to skip the model fitting because it's already been done, set `fit.multinomial <- FALSE`. \cr

One thing we do here that warrants mentioning is that we take all R-squared values < a user-defined value (`min.R2`) and assign them value `min.R2`. This is done because occassionally a dataset will produce an very negataive R2 value (e.g. -100) and we do not want these rare outlier values to have strong influence on the model fitting process. The default value for `min.R2` is -1.


```{r, eval=TRUE, results="hide"}
  
  model.file <- "nnet_mod_R2_P"
  moddir <- system.file("models", package="Stickbreaker")
  modpath <- paste(moddir, model.file, sep="/")
  data.file <- "Training_simulated_priors_fit_data"

  fit.multinomial <- TRUE
  min.R2 <- -1
  
  if (fit.multinomial == TRUE){
  
    mod.formula <- as.formula(model ~ R2.stick + R2.mult + R2.add + P.stick + P.mult + P.add)
    muts.to.sim <- c(3,4,5)
    fit.nmet.models <- vector("list", length(muts.to.sim))
    for (mut.i in 1:length(muts.to.sim)){
      n.muts <- muts.to.sim[mut.i]
      indir <- system.file("extdata", package="Stickbreaker")
      file.name <- paste(data.file, "_", n.muts, "muts.txt", sep="")
      inpath <- paste(indir, file.name, sep="/")
      if (file.exists(inpath)){
        data <- read.table(file=inpath, header=TRUE)
        data$R2.stick[which(data$R2.stick < min.R2)] <- min.R2
        data$R2.mult[which(data$R2.mult < min.R2)] <- min.R2
        data$R2.add[which(data$R2.add < min.R2)] <- min.R2
        fit.nmet.models[[mut.i]] <- fit.nnet.multinomial.regression(data, mod.formula)
        m <- fit.nnet.multinomial.regression(data, mod.formula)
        modelout <- paste(modpath, "_", n.muts, "muts.rda", sep="")
        saveRDS(m, file=modelout)
      }
    }  
  }
```

### 2. Incomplete networks {-}

This is very similar to the full networks, except that we supply a particular dataset structure (in the form of `geno.matrix`) and the simualtions are tailored to it. 

#### a. Generate a partial dataset to analyze

To have a dataset to work with, let's generate one. Suppose we have five mutations and want only engineer the first and last steps (i.e. wildtype plus each indivdual mutation and 5-mutant genotypes with each indivdiaul mutation removed.) 

```{r}
  n.muts <- 5
  coes <- 0.2
  sigma <- 0.03
  geno.matrix <- generate.geno.matrix(5)
  geno.matrix <- geno.matrix[which(rowSums(geno.matrix) %in% c(0,1,4,5)),]
  fit.matrix <- simulate.stick.data(n.muts, coes=coes, sigma=sigma, d.true=1, w.wt=1, geno.matrix)
```

#### b. Simulate data with same structure from priors {-}

Just as with full data we now need to simulate data from the priors. However, we want every dataset to have the same structure as the data we just cooked up (i.e. first and last steps only).

```{r, eval=FALSE}
  outdir <- system.file("extdata", package="Stickbreaker")
  file.name <- paste("Training_simulated_priors_fit_data_first_last", n.muts, "muts.txt", sep="")
  simdata.outpath <- paste(outdir, file.name, sep="/")
  coes.prior <- c(0.05, 0.5)
  sig.prior <- c(0, 0.25)
  n.samps.per.mod <- 10
  d.true <- 1
  d.range <- c(0.1, 10)
  d.adj.max <- 1.1
  w.wt <- 1
  wts <- c(2,1)
  n.samps.per.mod <- 100

  simulate.partial.data.from.priors.for.mod.selection(geno.matrix,
                                                      coes.prior=coes.prior,
                                                      sigs.prior=sig.prior,
                                                      mods.to.sim=c("stick", "mult", "add"),
                                                      d.true=1,
                                                      d.range=d.range,
                                                      d.adj.max=d.adj.max,
                                                      w.wt=1,
                                                      wts=wts,
                                                      outpath=simdata.outpath,
                                                      n.samps.per.mod=n.samps.per.mod,
                                                      coe.sim.model="identical",
                                                      coe.dist.par=NA,
                                                      print.interval=25)

```

#### c. Fit training data to multinomial model for model selection {-}

Everything here is exactly the same as with full sets of data.


```{r, eval=TRUE, results="hide"}
  
  model.file <- "nnet_mod_R2_P_5muts_first_last"
  moddir <- system.file("models", package="Stickbreaker")
  modpath <- paste(moddir, model.file, sep="/")

  fit.multinomial <- TRUE
  min.R2 <- -1
  
  if (fit.multinomial == TRUE){
    mod.formula <- as.formula(model ~ R2.stick + R2.mult + R2.add + P.stick + P.mult + P.add)
    indir <- system.file("extdata", package="Stickbreaker")
    file.name <- paste("Training_simulated_priors_fit_data_first_last", n.muts, "muts.txt", sep="")
    inpath <- paste(indir, file.name, sep="/")
      if (file.exists(inpath)){
        data <- read.table(file=inpath, header=TRUE)
        data$R2.stick[which(data$R2.stick < min.R2)] <- min.R2
        data$R2.mult[which(data$R2.mult < min.R2)] <- min.R2
        data$R2.add[which(data$R2.add < min.R2)] <- min.R2
        m <- fit.nnet.multinomial.regression(data, mod.formula)
        modelout <- paste(modpath, "_", n.muts, "muts.rda", sep="")
        saveRDS(m, file=modelout)
      }
    }  
```

## Model Selection Assessment {-}

A line at the end.
